Data preparation for employee category analysis

### 2023-04-25 04. Analysis/scripts/02_data_preparation_clusters_employee.Rmd
Update 02_data_preparation_clusters_employee.Rmd
Completing co-location metrics, starting rule-based groups

```{r}
prep_a <- 
  cohort_panel %>% 
  filter(daily_status_imputed)
  select
```

#### Proportion of working days co-located with supervisor
#### Proportion of working days co-located with entire team
#### Proportion of working days entire team is remote

>>>>

Identify employees who report to the site and collect `person_id`, `dates`, `location`, and `supervisor_e_mail_address`.

```{r}
prep_a <-
  cohort_panel %>%
  filter(daily_status_imputed == "Reporting to Job Site") %>%
  select(person_id, date, location, supervisor_e_mail_address) %>%
  distinct()
```

Self left join on `dates`, `location`, and `supervisor_e_mail_address` and filter to rows where the `person_id`s do not match.

```{r}
prep_b <- 
  prep_a %>% 
  left_join(prep_a,
            by = c("date", "location", "supervisor_e_mail_address"),
            suffix = c("","_joined")) %>%
  filter(person_id != person_id_joined)
```

Count dates where co-location occurs by `person_id`.

```{r}
prep_c <- 
  prep_b %>% 
  group_by(person_id) %>% 
  summarise(days_colocated_1 = n_distinct(date)) %>% 
  ungroup()
```

```{r}
prep_c %>% 
  skim()
```

Full join `working_days` from `prep_building`, calculate proportion of co-located days with at least one team member by `working_days` and `observation_window_days`, replace nulls with 0, and remove `working_days`.

```{r}
prep_xx <- 
  prep_c %>% 
  full_join(prep_building %>% 
              select(person_id,working_days)) %>% 
  mutate(days_colocated_1_working_prop = days_colocated_1/working_days,
         days_colocated_1_observation_prop = days_colocated_1/observation_window_days) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  select(-working_days)
```

#### Proportion of working days co-located with supervisor

Identify employees who report to the site and collect `person_id`, `dates`, `location`, `employee_e_mail_address`, and `supervisor_e_mail_address`.

```{r}
prep_a <-
  cohort_panel %>%
  filter(daily_status_imputed == "Reporting to Job Site") %>%
  select(person_id, date, location, employee_e_mail_address, supervisor_e_mail_address) %>%
  distinct()
```

Self inner join on `dates`, `location`, and `supervisor_e_mail_address` equal to `employee_e_mail_address` in the right table, and filter to rows where the `person_id`s do not match.

```{r}
prep_b <- 
  prep_a %>% 
  inner_join(prep_a,
            by = c("date", "location", "supervisor_e_mail_address" = "employee_e_mail_address"),
            suffix = c("","_joined")) 
```

Count dates where co-location with supervisor occurs by `person_id`.

```{r}
prep_c <- 
  prep_b %>% 
  group_by(person_id) %>% 
  summarise(days_colocated_supervisor = n_distinct(date)) %>% 
  ungroup()
```

```{r}
prep_c %>% 
  skim()
```

Full join `working_days` from `prep_building`, calculate proportion of co-located days with at least one team member by `working_days` and `observation_window_days`, replace nulls with 0, and remove `working_days`.

```{r}
prep_xx <- 
  prep_c %>% 
  full_join(prep_building %>% 
              select(person_id,working_days)) %>% 
  mutate(days_colocated_supervisor_working_prop = days_colocated_supervisor/working_days,
         days_colocated_supervisor_observation_prop = days_colocated_supervisor/observation_window_days) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  select(-working_days)
```

#### Proportion of working days co-located with entire team

Identify employees who report to the site and collect `person_id`, `dates`, `location`, and `supervisor_e_mail_address`.

```{r}
prep_a <-
  cohort_panel %>%
  filter(daily_status_imputed == "Reporting to Job Site") %>%
  select(person_id, date, location, supervisor_e_mail_address) %>%
  distinct()
```

Self left join on `dates`, `location`, and `supervisor_e_mail_address`.

```{r}
prep_b <- 
  prep_a %>% 
  left_join(prep_a,
            by = c("date", "location", "supervisor_e_mail_address"),
            suffix = c("","_joined"))
```

Count number of team members co-located by `person_id` and `date` by counting unique `person_id_joined`.

```{r}
prep_c <- 
  prep_b %>% 
  group_by(person_id,date,supervisor_e_mail_address) %>% 
  summarise(colocated_team_members = n_distinct(person_id_joined)) %>% 
  ungroup()
```

```{r}
prep_c %>% 
  skim()
```

```{r}
prep_d <- 
  cohort_panel %>% 
  group_by(date,supervisor_e_mail_address) %>% 
  summarise(team_size = n_distinct(person_id)) %>% 
  ungroup()
```

```{r}
prep_d %>% 
  skim()
```

Left join co-located team members and teams size, but `date` and `supervisor_e_mail_address`, then filter to when co-located team members is the same size as the team

```{r}
prep_e <- 
  prep_c %>% 
  left_join(prep_d) %>% 
  filter(colocated_team_members == team_size)
```

```{r}
prep_e %>% 
  skim()
```

Count dates where co-location with whole team occurs by `person_id`.

```{r}
prep_f <- 
  prep_e %>% 
  group_by(person_id) %>% 
  summarise(days_colocated_team = n_distinct(date)) %>% 
  ungroup()
```

```{r}
prep_f %>% 
  skim()
```

Full join `working_days` from `prep_building`, calculate proportion of co-located days with team member by `working_days` and `observation_window_days`, replace nulls with 0, and remove `working_days`.

```{r}
prep_xx <- 
  prep_f %>% 
  full_join(prep_building %>% 
              select(person_id,working_days)) %>% 
  mutate(days_colocated_team_working_prop = days_colocated_team/working_days,
         days_colocated_team_observation_prop = days_colocated_team/observation_window_days) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  select(-working_days)
```

#### Proportion of working days entire team is remote

Identify employees who telework and collect `person_id`, `dates`, `location`, and `supervisor_e_mail_address`.

```{r}
prep_a <-
  cohort_panel %>% 
  filter(daily_status_imputed == "Teleworking") %>%
  select(person_id, date, location, supervisor_e_mail_address) %>%
  distinct()
```

Self left join on `dates`, `location`, and `supervisor_e_mail_address`.

```{r}
prep_b <- 
  prep_a %>% 
  left_join(prep_a,
            by = c("date", "location", "supervisor_e_mail_address"),
            suffix = c("","_joined"))
```

Count number of team members co-located by `person_id` and `date` by counting unique `person_id_joined`.

```{r}
prep_c <- 
  prep_b %>% 
  group_by(person_id,date,supervisor_e_mail_address) %>% 
  summarise(telework_team_members = n_distinct(person_id_joined)) %>% 
  ungroup()
```

```{r}
prep_c %>% 
  skim()
```

```{r}
prep_d <- 
  cohort_panel %>% 
  group_by(date,supervisor_e_mail_address) %>% 
  summarise(team_size = n_distinct(person_id)) %>% 
  ungroup()
```

```{r}
prep_d %>% 
  skim()
```

Left join telework team members and teams size, but `date` and `supervisor_e_mail_address`, then filter to when telework team members is the same size as the team

```{r}
prep_e <- 
  prep_c %>% 
  left_join(prep_d) %>% 
  filter(telework_team_members == team_size)
```

```{r}
prep_e %>% 
  skim()
```

Count dates where telework with whole team occurs by `person_id`.

```{r}
prep_f <- 
  prep_e %>% 
  group_by(person_id) %>% 
  summarise(days_telework_team = n_distinct(date)) %>% 
  ungroup()
```

```{r}
prep_f %>% 
  skim()
```

Full join `working_days` from `prep_building`, calculate proportion of telework days with team member by `working_days` and `observation_window_days`, replace nulls with 0, and remove `working_days`.

```{r}
prep_xx <- 
  prep_f %>% 
  full_join(prep_building %>% 
              select(person_id,working_days)) %>% 
  mutate(days_telework_team_working_prop = days_telework_team/working_days,
         days_telework_team_observation_prop = days_telework_team/observation_window_days) %>% 
  mutate(
    across(
      .cols = where(is.numeric),
      .fns = ~replace_na(.,0)
    )
  ) %>% 
  select(-working_days)
```

## Define rule-based groups

The analysis plan defines nine rule-based groups for analysis:

1.  **All remote** - A policy definition for staff that are 100% remote
2.  **Teleworker** - A policy definition for staff that require two days
    per pay period onsite.
3.  **Onsite Flexible** - Defined by `position_category_description` in
    the roster data
4.  **Offsite** - Defined by `position_category_description` in the
    roster data
5.  **Onsite Required** - Defined by `position_category_description` in
    the roster data
6.  **Primarily remote** - Employees who work remotely above the 95th
    percentile <!-- could be 75th -->
7.  **Mostly remote** - Employees who work remotely between the median
    and the 95th percentile <!-- could be 75th -->
8.  **Mostly onsite** - Employees who work remotely between 5th
    percentile and the median <!-- could be 25th -->
9.  **Primarily onsite** - Employees who work remotely below the 5th
    percentile <!-- could be 75th -->

Below the variables for the groups are created.

### All remote

Define workers with 100% of their time as **all remote**.

```{r}
prep_xx <- 
  prep_building %>% 
  mutate(
    rule_based_all_remote = case_when(
      daily_status_teleworking_working_prop == 1 ~ 1,
                                      TRUE ~ 0)
  ) %>% 
  select(person_id,starts_with("rule_based_"))
```

```{r}
prep_building <- 
  prep_building %>% 
  left_join(prep_xx)
```
### Teleworker

First define pay periods based on this: https://www.gsa.gov/cdnstatic/GSA_Payroll_Calendar_2022_Revised.pdf. File is also saved in `r file.path(params$google_drive_home,"04. Analysis","documents")`.

```{r}
prep_a <- 
  tibble(
  date = seq.Date(from = ymd("2022-01-01"),
         to = ymd("2022-12-31"),
         by = "2 weeks")
) %>% 
  rownames_to_column(var = "period_number") 
# %>% 
#   mutate(
#     date_pay_period_start_lead = lead(date_pay_period_start)
  # )
```

```{r}
prep_b <- 
  tibble(
  date = seq.Date(from = ymd("2022-01-01"),
         to = ymd("2022-12-31"),
         by = "days")
)
```

```{r}
prep_c <- 
  prep_b %>% 
  left_join(prep_a) %>% 
  fill(period_number, .direction = "down")
```

```{r}
prep_d <- 
  cohort_panel %>% 
  select(person_id,date, daily_status_imputed) %>% 
  left_join(prep_c)
```

```{r}
prep_e <- 
      prep_d %>% 
  group_by(person_id,daily_status_imputed,period_number) %>% 
  summarise(days = n_distinct(date)) %>% 
  ungroup()
```

#CONTINUE FROM HERE

```{r}
prep_e %>% 
  filter(daily_status_imputed == "Reporting to Job Site") %>% 
  mutate(
    telework_onsite_minimum = case_when(days >= 2 ~ 1,
                                        TRUE ~ 0)
  ) %>% 
  group_by(person_id) %>% 
  summarise(
    telework_onsite_minimum = mean(telework_onsite_minimum)
  )
```

### Primarily, mostly onsite and remote

First, calculate the quantiles for probabilities 0.50 and 0.95.

```{r}
prep_a <- 
  prep_building %>% 
  summarise(
    across(
      .cols = c(daily_status_reporting_to_job_site_working_prop,daily_status_teleworking_working_prop,
                daily_status_reporting_to_job_site_observation_prop,daily_status_teleworking_observation_prop),
      .fns = list(q_95 = ~quantile(.,probs = .95),
                  q_50 = ~quantile(.,probs = .50))
    )
  )
```

Use the quantiles above to create new variables for `primarily_` and `mostly_` working remote or onsite.

```{r}
prep_xx <-
  prep_building %>%
  select(
    person_id,
    daily_status_reporting_to_job_site_working_prop,
    daily_status_teleworking_working_prop,
    daily_status_reporting_to_job_site_observation_prop,
    daily_status_teleworking_observation_prop
  ) %>%
  mutate(
    rule_based_primarily_reporting_to_job_site_working =
      case_when(
        daily_status_reporting_to_job_site_working_prop >= prep_a$daily_status_reporting_to_job_site_working_prop_q_95 ~ 1,
        TRUE ~ 0
      ),
    rule_based_primarily_teleworking_working =
      case_when(
        daily_status_teleworking_working_prop >= prep_a$daily_status_teleworking_working_prop_q_95 ~ 1,
        TRUE ~ 0
      ),
    rule_based_primarily_reporting_to_job_site_observation =
      case_when(
        daily_status_reporting_to_job_site_observation_prop >= prep_a$daily_status_reporting_to_job_site_observation_prop_q_95 ~ 1,
        TRUE ~ 0
      ),
    rule_based_primarily_teleworking_observation =
      case_when(
        daily_status_teleworking_observation_prop >= prep_a$daily_status_teleworking_observation_prop_q_95 ~ 1,
        TRUE ~ 0
      ),
    rule_based_mostly_reporting_to_job_site_working =
      case_when(
        daily_status_reporting_to_job_site_working_prop >= prep_a$daily_status_reporting_to_job_site_working_prop_q_50 ~ 1,
        TRUE ~ 0
      ),
    rule_based_mostly_teleworking_working =
      case_when(
        daily_status_teleworking_working_prop >= prep_a$daily_status_teleworking_working_prop_q_50 ~ 1,
        TRUE ~ 0
      ),
    rule_based_mostly_reporting_to_job_site_observation =
      case_when(
        daily_status_reporting_to_job_site_observation_prop >= prep_a$daily_status_reporting_to_job_site_observation_prop_q_50 ~ 1,
        TRUE ~ 0
      ),
    rule_based_mostly_teleworking_observation =
      case_when(
        daily_status_teleworking_observation_prop >= prep_a$daily_status_teleworking_observation_prop_q_50 ~ 1,
        TRUE ~ 0
      ),
  ) %>%
  select(person_id, starts_with("rule_based_"))
```

#CONTINUE FROM HERE
